High-resolution ALMA observations of V4046\,Sgr: a circumbinary disc with a thin ring

### 2020-10-16 Draft_V4046Sgr_mnras.tex
Update Draft_V4046Sgr_mnras.tex

% Abstract of the paper
\begin{abstract}
  The nearby V4046\,Sgr spectroscopic binary hosts a gas-rich disc known for its wide cavity and dusty ring. We present new high resolution ($\sim$20\, mas) ALMA observations of the 1.3\,mm continuum. The comparison of these observations, combined with SPHERE-IRDIS polarized images and a well-sampled spectral energy distribution (SED), against radiative transfer (RT) predictions carried out with the RADMC3D package, allow us to propose a physical model for the source. The new ALMA data reveal a very fine ring at a radius of 13.46$\pm$0.43\,au (Ring13), with a marginally resolved radial width of 4.17$\pm$0.94\,au. Ring13 is the brightest structure in scattered-light, where it is surrounded by a $\sim$9\,au-wide gap , and it is flanked by a very bright mm outer ring (Ring24) with a sharp inner edge at 24\,au. The steeply decreasing radial slope of Ring24 breaks at $\sim$35\,au into a shallow tail. The RT model requires an inner ring at $\sim$6\,au (Ring6) in small dust grains, hiding under the IRDIS coronagraph, and that surrounds an inner circumbinary disk. Faint mm-continuum coincident with Ring6 is picked up by ALMA, whose morphology suggests that Ring6 is lopsided or shadowed by the secondary. The previously reported scattered-light shadow of the secondary star is also reproduced by the RT model. The surpising thin Ring13 is nonetheless **15** times wider than the model scale height, and could thus be long-lived. The strong near-far disc asymmetry at 1.65\,$\micron$ points at a very forward-scattering phase function, and requires grain radii of no less than 0.4\,$\micron$. 
\end{abstract}

@@

Recent observations of young circumstellar discs have transformed current knowledge of planet formation, but the focus of resolved imaging with the Atacama Large Millimeter/submillimetre Array (ALMA) or with VLT/SPHERE has mainly been towards the brighter sources \citep{Andrews2020arXiv200105007A}. The Disks ARound T Tauri Stars (DARTTS) program was first presented in \citet{Avenhaus_2018} with the aim to image eight stars with the Spectro-Polarimeter High-contrast Exoplanet REsearch (SPHERE). The sample is not biased towards exceptionally bright and large disks, and consists of only solar-mass stars. A second part of the survey increased the number of sources by presenting 21 new images of circumstellar disks \citep{Garufi2020}. The observations revealed diversed structures and morphologies in the scattering surface of these disks. This letter on V4046 Sagittarii (Sgr) is part of our new DARTTS survey with ALMA (DARTTS-A) which will present millimeter observations of nine protoplanetary disks previously imaged in polarized scattered light in DARTTS-S.

@@

Repeating the optimization of the disk orientation, but this time aiming for Ring24 with a radial domain from 20\,au to 70\,au, we obtained a PA of 76.78$\pm$0.03\,$\degr$, with an inclination of 34.01$\pm$0.01\,$\degr$ and a centre at $\Delta \alpha = 19\pm0.04$\,mas $\Delta \delta = 7\pm0.03$\,mas relative to the stars. We see that both Ring13 and Ring24 both share / DO NOT SHARE? the same orientation and centre, given the errors, and both are offset from the star. Even though all orientation parameters for Ring13 and Ring24 are consistent within the errors, there are however some hints for a different orientation, as summarised in Fig. \ref{fig:polarring}, perhaps due to the joint effect of all these small differences. ****CALCULATE THE CHI2 DIFFERENCE BETWEEN THE TWO TRACES? SOME FORM OF STATS WOULD BE GOOD****

***I WOULD REMOVE THIS OR IDENTIFY SOMETHING CONCRETE OUT OF IT: We can compare the measured location of the inner ring according to these parameters versus to the location obtained before. The deviation for each azimuth is shown in Fig. \ref{fig:polarring}, and is not trivial since it depend on a few parameters such as the different inclinations of the rings, the difference in position angle and the shift of the centres relative to the stars. But even though those parameters play an important role, a possible small difference in eccentricities between the inner and the outer ring could also help explain this deviation between the two measurements. ***

>>>>

% Abstract of the paper
\begin{abstract}
  The nearby V4046\,Sgr spectroscopic binary hosts a gas-rich disc known for its wide cavity and dusty ring. We present new high resolution ($\sim$40\,mas) ALMA observations of the 1.3\,mm continuum. The comparison of these observations, combined with SPHERE-IRDIS polarized images and a well-sampled spectral energy distribution (SED), against radiative transfer (RT) predictions carried out with the RADMC3D package, allow us to propose a physical model for the source. The new ALMA data reveal a very fine ring at a radius of 13.46$\pm$0.43\,au (Ring13), with a marginally resolved radial width of 4.17$\pm$0.94\,au. Ring13 is the brightest structure in scattered-light, where it is surrounded by a $\sim$9\,au-wide gap, and it is flanked by a very bright mm outer ring (Ring24) with a sharp inner edge at 24\,au. The steeply decreasing radial slope of Ring24 breaks at $\sim$35\,au into a shallow tail. The RT model requires an inner ring at $\sim$6\,au (Ring6) in small dust grains, hiding under the IRDIS coronagraph, and that surrounds an inner circumbinary disk. Faint mm-continuum coincident with Ring6 is picked up by ALMA, whose morphology suggests that Ring6 is lopsided or shadowed by the secondary. The previously reported scattered-light shadow of the secondary star is also reproduced by the RT model. The surprising thin Ring13 is nonetheless 15 times wider than the model scale height, and could thus be long-lived. The strong near-far disc asymmetry at 1.65\,$\micron$ points at a very forward-scattering phase function, and requires grain radii of no less than 0.4\,$\micron$. 
\end{abstract}

@@

Recent observations of young circumstellar discs have transformed current knowledge of planet formation, but the focus of resolved imaging with the Atacama Large Millimeter/submillimetre Array (ALMA) or with VLT/SPHERE has mainly been towards the brighter sources \citep{Andrews2020arXiv200105007A}. The Disks around T Tauri Stars (DARTTS) program was first presented in \citet{Avenhaus_2018} with the aim to image eight stars with the Spectro-Polarimeter High-contrast Exoplanet REsearch (SPHERE). The sample is not biased towards exceptionally bright and large disks, and consists of only solar-mass stars. A second part of the survey increased the number of sources by presenting 21 new images of circumstellar disks \citep{Garufi2020}. The observations revealed diverse structures and morphologies in the scattering surface of these disks. This letter on V4046 Sagittarii (Sgr) is part of our new DARTTS survey with ALMA (DARTTS-A) which will present millimetre observations of nine protoplanetary disks previously imaged in polarized scattered light in DARTTS-S.

@@

Repeating the optimization of the disk orientation, but this time aiming for Ring24 with a radial domain from 20\,au to 70\,au, we obtained a PA of 76.78$\pm$0.03\,$\degr$, with an inclination of 34.01$\pm$0.01\,$\degr$ and a centre at $\Delta \alpha = 19\pm0.04$\,mas $\Delta \delta = 7\pm0.03$\,mas relative to the stars. We see that both Ring13 and Ring24 both share the same orientation and centre, given the errors, and both are offset from the star. Even though all orientation parameters for Ring13 and Ring24 are consistent within the errors, there are however some hints for a different orientation, as summarised in Fig. \ref{fig:polarring}, perhaps due to the joint effect of all these small differences. ****CALCULATE THE CHI2 DIFFERENCE BETWEEN THE TWO TRACES? SOME FORM OF STATS WOULD BE GOOD****
